# AlphaBash

You can do amazing things with bash. For example, let's pick a dataset. ...

```{r Pick-a-dataset}
require(tidyverse)
# There's some here https://vincentarelbundock.github.io/Rdatasets/articles/data.html
install.packages("AER")
require(AER)
data(Fatalities)
write_tsv(Fatalities,"data/fatalaties.tsv")
```

Now, let's use the awk language, named after its three creators Alfred Aho, Brian Kernighan, and Peter J. Weinberger around 1977. Note, the quarto chunk has changed language.

I'm going to ask awk to print the first column of this dataframe, followed by a tab, followed by the filename. This could be useful when you want to extract a number of columns from several files and include the filename.

```{bash Awk}
awk '{print $1"\t"FILENAME}' data/fatalaties.tsv
```

# Analysing the data

Once you've cleaned the data, it can be loaded into R and sorted using R verbs and plotted using ggplot2.

```{r Load-data}
x <- read_tsv("data/fatalaties.tsv")

y <- x %>% filter(state!="al") %>% 
  rename(State=state) %>% 
  mutate(State_year=paste(State,year,sep="_")) %>% 
  select(-c(State_year)) %>% 
  filter(!grepl("z",State)) %>% 
  mutate(New=gsub("1","One",spirits))
```

```{r}
require(ggsci)
require(ggprism)
z <- y %>% ggplot(aes(x=spirits, 
                      y=income, 
                      colour=State)) +
  geom_point() +
  scale_colour_lancet()
z
```


```{r}
ggsave(z,file="plot.pdf",height=4,width=4)
```

